### 2016-08-18 latex/paper.tex
Fix figure order
Add term list

@@

...

@@

\begin{figure}[!h]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-06.eps}
\caption{caption}
\label{fig:schematic}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-07.eps}
\caption{Deformation of a rim cross-section. The shear center $S\rightarrow S'$ translates by the vector $\bm{u} = u\hat{\bm{e}}_1 + v\hat{\bm{e}}_2 + w\hat{\bm{e}}_3$ and rotates through an angle $\phi$.}
\label{fig:def}
\end{figure}

@@

In this paper we consider bicycle wheels with slender prestressed spokes. We model a single spoke as an elastic bar pinned at each end. This ensures that the spoke force is always collinear with the spoke axis. We fix our global coordinate system to the hub, which is assumed to be rigid. Therefore, the displacement of the hub node of a spoke is zero. The linearized elongation of a spoke is given by
	\begin{equation}\label{eq:selong}
	\Delta l = \bm{u}_n\cdot \hat{\bm{n}}_1
	\end{equation}
where $\bm{u}_n$ is the displacement vector of the spoke nipple and $\hat{\bm{n}}_1$ is a unit vector pointing from the spoke nipple to the hub eyelet. The linearized rotation of the spoke is given by
	\begin{equation}\label{eq:srot1}
	\Omega_1 = \frac{1}{l} \bm{u}_n\cdot \hat{\bm{n}}_2
	\end{equation}
	\begin{equation}\label{eq:srot2}
	\Omega_2 = \frac{1}{l} \bm{u}_n\cdot \hat{\bm{n}}_3
	\end{equation}
where $\hat{\bm{n}}_2$ and $\hat{\bm{n}}_3$ are two orthogonal unit vectors orthogonal to $\hat{\bm{n}}_1$, and $l$ is the spoke length.

The elongation produces a net force on the rim parallel to the original spoke axis equal to
	\begin{equation}\label{eq:sF1}
	f_1 = E_sA_s\left(\frac{\Delta l}{l}\right)\hat{\bm{n}}_1 = \frac{E_sA_s}{l} \hat{\bm{n}}_1 \cdot \bm{u}_n 
	\end{equation}
The rotation of the spoke produces net forces on the rim perpendicular to the original spoke axis due to the rotation of the initial tension vector.
	\begin{equation}\label{eq:sF2}
	f_2 = T \sin{\Omega_1} \approx \frac{T}{l} \hat{\bm{n}}_2 \cdot \bm{u}_n
	\end{equation}
	\begin{equation}\label{eq:sF3}
	f_3 = T \sin{\Omega_2} \approx \frac{T}{l} \hat{\bm{n}}_3 \cdot \bm{u}_n
	\end{equation}
since $\hat{\bm{n}}_1,\hat{\bm{n}}_2,\hat{\bm{n}}_3$ are mutually orthogonal, the total force on the rim is given by $\bm{f} = \bm{k}_f \cdot \bm{u}_n$, where the spoke stiffness matrix is given by
	\begin{equation}\label{eq:kf}
	\bm{k}_f = \frac{E_sA_s}{l}\hat{\bm{n}}_1\hat{\bm{n}}_1 + \frac{T}{l}(\hat{\bm{n}}_2\hat{\bm{n}}_2 + \hat{\bm{n}}_3\hat{\bm{n}}_3)
	\end{equation}
The first term in Eqn.~\ref{eq:kf} is the {\bf elastic stiffness} and the second term is the {\bf tension stiffness}. The tension stiffness is the component responsible for the transverse vibrations of a guitar string, e.g.

Following the approach of Smith~\cite{Smith1901a} and Pippard~\cite{Pippard1932d}, we will approximate the spoke system as a continuum which exerts a distributed load on the rim. The stiffness per unit length is defined by the sum of the spoke stiffnesses in one periodic grouping of spokes, divided by the arc length of the grouping.
	\begin{equation}\label{eq:kbar}
	\bar{\bm{k}} = \frac{1}{2\pi R}\left(\frac{n_s}{n_p}\right) \sum_i^{n_p} \bm{k}_{f, i}
	\end{equation}
	The distributed load $\bar{\bm{f}}$ does an amount of work per unit length equal to
	\begin{equation}\label{eq:w_spokes}
	\bar{w} = \int_0^{u_n} \bar{\bm{f}} \cdot d\bm{u}_n = \frac{1}{2} \bm{u}_n \cdot \bar{\bm{k}} \cdot \bm{u}_n
	\end{equation}
The line of action is assumed to pass sufficiently close to the rim shear center that any difference between the displacement of the spoke nipple and the displacement of the shear center can be ignored. Therefore $\bm{u}_n = \bm{u}$. Substituting a lateral buckling displacement $\bm{u}=u\hat{\bm{e}}_1$ into Eqn. \ref{eq:w_spokes} and integrating over the length of the rim gives the strain energy stored in the spoke system.
	\begin{equation}\label{eq:Us}
	U_{spokes} = \int_0^{2\pi} \bar{w} \, ds = \frac{1}{2} \bar{k}_{uu}u^2
	\end{equation}
The continuum stiffness can be written $\bar{\bm{k}} = \bar{\bm{k}}^{el} + T\bar{\bm{k}}^{tens}$, where the dependence on $T$ is shown explicitly. A very close approximation for the $\bar{k}_{uu}$ component for a symmetrically dished wheel is given by
	\begin{equation}\label{eq:kuu}
	\bar{k}_{uu} = \frac{n_sE_sA_s}{2\pi Rl}\sin^2{\alpha} + \frac{n_s T}{2\pi Rl}
	\end{equation}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-05.eps}
\caption{\textbf{(a)} Buckling mode parameter map. Only the first 4 modes are shown. \textbf{(b)}--\textbf{(c)} Normalized buckling tension vs. $\tilde{\mu}$ and $\lambda$. Solid black line = Eqn. \ref{eq:Tc_nondim}, dashed red line = power law, blue stars = finite-element simulations. In (a) $\lambda=10$ while in (b) $\tilde{\mu} = 0.38$.}
\label{fig:Tc_nondim}
\end{figure}

@@

\subsection{Rims with low torsional stiffness}\label{sec:powerlaw_1}
Many bicycle rims are built from thin-walled open channels which have very low torsional stiffness compared with their bending stiffness. Replacing the discrete mode number $n$ with a continuous variable $\bar{n}$, we adopt the following ansatz: (a) $\bar{n}^2 \gg 1$ and (b) $\tilde{\mu}\bar{n}^2 \ll 1$. Under these conditions, Eqn. \ref{eq:Tc_nondim} becomes
	\begin{equation}\label{eq:Tcn_small_mu}
	T_{c,n} = \frac{2\pi EI_{11}}{n_sR^2} \left( \tilde{\mu}\bar{n}^4 + \frac{\lambda}{\bar{n}^2}\right)
	\end{equation}
Minimizing Eqn. \ref{eq:Tcn_small_mu} with respect to $\bar{n}$ by setting the derivative $dT_{c,n}/d\bar{n}=0$, we obtain the scaling law
	\begin{equation}\label{eq:nbar_small_mu}
	\bar{n} = \left(\frac{\lambda}{2\tilde{\mu}} \right)
	\end{equation}
Substituting Eqn. \ref{eq:nbar_small_mu} into Eqn. \ref{eq:Tc_small_mu} yields, after simplification, a simple power law relationship.
	\begin{equation}\label{eq:Tc_small_mu}
	T_c \approx \frac{11.875}{n_sR^2} \, \left(\bar{k}_{uu}^{el}R^4 \right)^{2/3} \widetilde{GJ}^{1/3}
	\end{equation}
Equation \ref{eq:Tc_small_mu} suggests that when the torsional stiffness is smaller than the bending stiffness, the buckling behavior is dominated by torsion and does not depend explicitly on the bending stiffness at all. Equation \ref{eq:Tc_small_mu} is shown in Fig. \ref{fig:Tc_nondim} next to Equation \ref{eq:Tc_nondim} and finite-element calculations. The power law is strictly less than Equation \ref{eq:Tc_nondim} which conveniently gives a conservative design criterion.

@@

>>>>

\newcommand{\e}{\ensuremath{\hat{\bm{e}}_1}}
\newcommand{\ee}{\ensuremath{\hat{\bm{e}}_2}}
\newcommand{\eee}{\ensuremath{\hat{\bm{e}}_3}}

\newcommand{\n}{\ensuremath{\hat{\bm{n}}_1}}
\newcommand{\nn}{\ensuremath{\hat{\bm{n}}_2}}
\newcommand{\nnn}{\ensuremath{\hat{\bm{n}}_3}}

@@

@@

\begin{figure}[!h]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-05.eps}
\caption{caption}
\label{fig:schematic}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-06.eps}
\caption{Deformation of a rim cross-section. The shear center $S\rightarrow S'$ translates by the vector $\bm{u} = u\e + v\ee + w\eee$ and rotates through an angle $\phi$.}
\label{fig:def}
\end{figure}

@@

In this paper we consider bicycle wheels with slender prestressed spokes. We model a single spoke as an elastic bar pinned at each end. This ensures that the spoke force is always collinear with the spoke axis. We fix our global coordinate system to the hub, which is assumed to be rigid. Therefore, the displacement of the hub node of a spoke is zero. The linearized elongation of a spoke is given by
	\begin{equation}\label{eq:selong}
	\Delta l = \bm{u}_n\cdot \n
	\end{equation}
where $\bm{u}_n$ is the displacement vector of the spoke nipple and \n{} is a unit vector pointing from the spoke nipple to the hub eyelet. The linearized rotation of the spoke is given by
	\begin{equation}\label{eq:srot1}
	\Omega_1 = \frac{1}{l} \bm{u}_n\cdot \nn
	\end{equation}
	\begin{equation}\label{eq:srot2}
	\Omega_2 = \frac{1}{l} \bm{u}_n\cdot \nnn
	\end{equation}
where \nn{} and \nnn{} are two orthogonal unit vectors orthogonal to \n{}, and $l$ is the spoke length.

The elongation produces a net force on the rim parallel to the original spoke axis equal to
	\begin{equation}\label{eq:sF1}
	f_1 = E_sA_s\left(\frac{\Delta l}{l}\right)\n = \frac{E_sA_s}{l} \n \cdot \bm{u}_n 
	\end{equation}
The rotation of the spoke produces net forces on the rim perpendicular to the original spoke axis due to the rotation of the initial tension vector.
	\begin{equation}\label{eq:sF2}
	f_2 = T \sin{\Omega_1} \approx \frac{T}{l} \nn \cdot \bm{u}_n
	\end{equation}
	\begin{equation}\label{eq:sF3}
	f_3 = T \sin{\Omega_2} \approx \frac{T}{l} \nn \cdot \bm{u}_n
	\end{equation}
since \n{}, \nn{}, \nnn{} are mutually orthogonal, the total force on the rim is given by $\bm{f} = \bm{k}_f \cdot \bm{u}_n$, where the spoke stiffness matrix is given by
	\begin{equation}\label{eq:kf}
	\bm{k}_f = \frac{E_sA_s}{l}\n\n + \frac{T}{l}(\nn\nn + \nnn\nnn)
	\end{equation}
The first term in Eqn.~\ref{eq:kf} is the {\bf elastic stiffness} and the second term is the {\bf tension stiffness}. The tension stiffness is the component responsible for the transverse vibrations of a guitar string, e.g.

Following the approach of Smith~\cite{Smith1901a} and Pippard~\cite{Pippard1932d}, we will approximate the spoke system as a continuum which exerts a distributed load on the rim. The stiffness per unit length is defined by the sum of the spoke stiffnesses in one periodic grouping of spokes, divided by the arc length of the grouping.
	\begin{equation}\label{eq:kbar}
	\bar{\bm{k}} = \frac{1}{2\pi R}\left(\frac{n_s}{n_p}\right) \sum_i^{n_p} \bm{k}_{f, i}
	\end{equation}
	The distributed load $\bar{\bm{f}}$ does an amount of work per unit length equal to
	\begin{equation}\label{eq:w_spokes}
	\bar{w} = \int_0^{u_n} \bar{\bm{f}} \cdot d\bm{u}_n = \frac{1}{2} \bm{u}_n \cdot \bar{\bm{k}} \cdot \bm{u}_n
	\end{equation}
The line of action is assumed to pass sufficiently close to the rim shear center that any difference between the displacement of the spoke nipple and the displacement of the shear center can be ignored. Therefore $\bm{u}_n = \bm{u}$. Substituting a lateral buckling displacement $\bm{u}=u\e$ into Eqn. \ref{eq:w_spokes} and integrating over the length of the rim gives the strain energy stored in the spoke system.
	\begin{equation}\label{eq:Us}
	U_{spokes} = \int_0^{2\pi} \bar{w} \, ds = \frac{1}{2} \bar{k}_{uu}u^2
	\end{equation}
The continuum spoke stiffness can be written $\bar{\bm{k}} = \bar{\bm{k}}^{el} + T\bar{\bm{k}}^{tens}$, where the dependence on $T$ is shown explicitly. A very close approximation for the $\bar{k}_{uu}$ component for a symmetrically dished wheel is given by
	\begin{equation}\label{eq:kuu}
	\bar{k}_{uu} = \frac{n_sE_sA_s}{2\pi Rl}\sin^2{\alpha} + \frac{n_s T}{2\pi Rl}
	\end{equation}

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=1.0]{figures/bmd_figures-04.eps}
\caption{\textbf{(a)} Buckling mode parameter map. Only the first 4 modes are shown. \textbf{(b)}--\textbf{(c)} Normalized buckling tension vs. $\tilde{\mu}$ and $\lambda$. Solid black line = Eqn. \ref{eq:Tc_nondim}, dashed red line = power law, blue stars = finite-element simulations. In (a) $\lambda=10$ while in (b) $\tilde{\mu} = 0.38$.}
\label{fig:Tc_nondim}
\end{figure}

@@

\subsection{Rims with low torsional stiffness}\label{sec:powerlaw_1}
Many bicycle rims are built from thin-walled open channels which have very low torsional stiffness compared with their bending stiffness. Replacing the discrete mode number $n$ with a continuous variable $\bar{n}$, we adopt the following ansatz: (a) $\bar{n}^2 \gg 1$ and (b) $\tilde{\mu}\bar{n}^2 \ll 1$. Under these conditions, Eqn. \ref{eq:Tc_nondim} becomes
	\begin{equation}\label{eq:Tcn_small_mu}
	T_{c,n} = \frac{2\pi EI_{11}}{n_sR^2} \left( \tilde{\mu}\bar{n}^4 + \frac{\lambda}{\bar{n}^2}\right)
	\end{equation}
Minimizing Eqn. \ref{eq:Tcn_small_mu} with respect to $\bar{n}$ by setting the derivative $dT_{c,n}/d\bar{n}=0$, we obtain the scaling law
	\begin{equation}\label{eq:nbar_small_mu}
	\bar{n} = \left(\frac{\lambda}{2\tilde{\mu}} \right)
	\end{equation}
Substituting Eqn. \ref{eq:nbar_small_mu} into Eqn. \ref{eq:Tc_small_mu} yields, after simplification, a simple power law relationship.
	\begin{equation}\label{eq:Tc_small_mu}
	T_c \approx \frac{11.875}{n_sR^2} \, \left(\bar{k}_{uu}^{el}R^4 \right)^{2/3} (\widetilde{GJ})^{1/3}
	\end{equation}
Equation \ref{eq:Tc_small_mu} suggests that when the torsional stiffness is smaller than the bending stiffness, the buckling behavior is dominated by torsion and does not depend explicitly on the bending stiffness at all. Equation \ref{eq:Tc_small_mu} is shown in Fig. \ref{fig:Tc_nondim} next to Equation \ref{eq:Tc_nondim} and finite-element calculations. The power law is strictly less than Equation \ref{eq:Tc_nondim} which conveniently gives a conservative design criterion.

@@

\section*{LIST OF TERMS}
The following notation is used in this paper:

\begin{tabular}{c|p{12cm}}
\hline
\textit{\textbf{Term}} & \textit{\textbf{Definition}} \\
\hline
$n_s$		& Number of spokes\\
$E_s, A_s$	& Spoke Young's modulus and cross-sectional area\\
$l$			& Spoke length\\
$\alpha$	& Out-of-plane spoke angle--wall angle of a hypothetical ``cone'' of spokes\\
$\beta$		& In-plane spoke angle--measured between the in-plane projection of the spoke and
			  a radius of the wheel\\
$\n$		& Unit vector aligned with the spoke axis\\
$\nn,\nnn$	& Two unit vectors perpendicular to the spoke axis\\
$\e,\ee,\eee$ & Coordinate axes defined at the rim shear center\\
$\bm{u}$	& Displacement of the rim shear center\\
$u,v,w$		& Vector components of $\bm{u}$ in the \e{}, \ee{}, \eee{} directions.\\
$\phi$		& Rotation of the rim cross-section about the $\hat{\bm{e}}_3$ axis.\\
$\delta u, \delta \phi$ & Virtual displacement and rotation\\
$u_n, \phi_n$ & Buckling mode coefficients\\
$n$			& Buckling mode number\\
$T$			& Tension in a spoke\\
$T_{cn}$	& Critical tension for the $n$th mode\\
$T_c$		& Critical buckling tension (minimum of $T_{cn}$ over all modes)\\
$T_e$		& ``Euler'' tension--tension which would produce an axial force in the rim
			  equal to the buckling load for a straight beam-column of length $2\pi R$\\
$\bm{f}$	& Force exerted by a spoke on the rim\\
$f_1, f_2,f_3$ & Vector components of $\bm{f}$\\
$\bm{k}_f$	& Stiffness matrix for a single spoke\\
$\bar{\bm{k}}$ & Equivalent stiffness per-unit-length of the spoke system. $\bar{\bm{k}} = \bar{\bm{k}}^{el} + \bar{\bm{k}}^{tens}$\\
$\bar{\bm{k}}^{el}$ & Part of $\bar{\bm{k}}$ proportional to $E_sA_s$\\
$\bar{\bm{k}}^{tens}$ & Part of $\bar{\bm{k}}$ proportional to $T$\\
$\bar{k}_{ij}$ & Tensor components of $\bar{\bm{k}}$\\
$R$			& Rim radius--measured from the rim shear center\\
$E$			& Rim Young's modulus\\
$I_{11}$	& Second moment of area of the rim in the lateral direction\\
$J$			& Rim torsion constant\\
$I_w$		& Rim warping constant\\
$\widetilde{GJ}$ & Effective torsional rigidity including both Saint-Venant torsion and warping\\
$\tilde{\mu}$ & Ratio of effective torsional rigidity to bending rigidity: $\widetilde{GJ}/EI_{11}$\\
$\lambda$	& Ratio of spoke system stiffness to rim bending stiffness OR ratio of rim radius to 
			  the characteristic length of a beam on an elastic foundation.\\
\hline
\end{tabular}
